Optimal Particle Paths Around Points in $\R^2$ with Constrained Acceleration

### 2013-09-29 cone_theorems.tex
Fixing trig in quadrilateral problem.
Reverted John's equations.
Finishing up the paper.

We would like to minimize the path length around the quadrilateral shown in Figure \ref{fig:quad}.

@@

\begin{align}
  x_c& = w - \frac{l_2}{2 \cos{\theta}}\\
  y_c& = \frac{l_2}{2} - \frac{l_2}{2 \sin{\theta}}\\
  l_{AB}& = \sqrt{x_c^2 + y_c^2}\\
  & = \sqrt{ \paren{\frac{l_1}{2} - \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }\\
  l_{BC}& = l_2\\
  l_{CD}& = \sqrt{\paren{\frac{l_1}{2} + \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }
\end{align}

Since we know that $l_{BC}$ is fixed the problem reduces to

@@

\begin{multline}
  \argmin_{\theta} \left[ 2\paren{\frac{l_1}{2} - \frac{l_2}{2 \sin\theta}}^2 \right.\\
  \left. + \paren{w - \frac{l_2}{2 \cos\theta}}^2 + \paren{w + \frac{l_2}{2 \cos\theta}}^2 \right]
\end{multline}

@@

>>>>

We would like to minimize the path length around the quadrilateral shown in Figure \ref{fig:quad}. Solving this problem will hopefully give us intuition for solving the problem of actual paths. We allow the quadrilateral's perimeter to vary based on an angle $\theta$ which has a physical representation of measuring how symmetric the system is.

@@

Thus, we want to choose the angle which will minimize the path traveled by a particle along the perimeter of the quadrilateral. Intuitively, a symmetric quadrilateral should provide the fastest route, but we will formalize this intuition. We do this by deriving the following equations using basic trigonometry:

\begin{align}
  x_c& = w - \frac{l_2}{2 \cos{\theta}}\\
  y_c& = \frac{l_2}{2} - \frac{l_2}{2} \sin{\theta}\\
  l_{AB}& = \sqrt{x_c^2 + y_c^2}\\
  & = \sqrt{ \paren{\frac{l_1}{2} - \frac{l_2}{2} \cos\theta}^2 + \paren{w - \frac{l_2}{2} \sin\theta}^2 }\\
  l_{BC}& = l_2\\
  l_{CD}& = \sqrt{\paren{\frac{l_1}{2} + \frac{l_2}{2} \cos\theta}^2 + \paren{w - \frac{l_2}{2} \sin\theta}^2 }
\end{align}

Since we know that $l_{BC}$ is fixed, our minimization problem reduces to

@@

\begin{multline}
  \argmin_{\theta} \left[ 2\paren{\frac{l_1}{2} - \frac{l_2}{2} \sin\theta}^2 \right.\\
  \left. + \paren{w - \frac{l_2}{2} \cos\theta}^2 + \paren{w + \frac{l_2}{2} \cos\theta}^2 \right]
\end{multline}

@@

Thus we see that the optimal angle for minimizing the path length is $\theta = \frac{\pi}{2}$. In other words, it is better to be have a symmetric path. This works well with our intuition, and so now we are ready to solve problems related to particles traveling around cones.

### 2013-09-29 path_theorems.tex
Finishing up the paper.

The simplest problem to tackle first, is the optimal trajectory from a current position, $\bvec{X_0}$ to a final position, $\bvec{X_f}$. Working in a polar coordinate system centered on $X_f$, it is clear, that our goal is achieved when $r = 0$. Furthermore, $\theta(t)$ must be $0$ before reaching $X_f$, otherwise, it will grow exponentially fast as $r(t) \to 0$, which can be seen in equation (\ref{eq:r-deriv}).

@@

\begin{theorem}
  Given a 2-dimensional particle $p$ whose initial position is $\bvec{X_1}$ and moves with bounded tangential acceleration, $a_t \le a_{t,max}$, and infinite centripetal acceleration, the fastest path $\hat{\gamma}(t)$ which $p$ can trace from $\bvec{X_1}$ to $\bvec{X_2}$ is the line segment from $\bvec{X_1}$ to $\bvec{X_2}$:
\end{theorem}

>>>>

Now that we have defined our notation, we are ready to think about the problem of optimal trajectories.

The simplest problem to tackle first, is the optimal trajectory from a starting position, $\bvec{X_0}$ to a final position, $\bvec{X_f}$. Working in a polar coordinate system centered on $X_f$, it is clear, that our goal is achieved when $r = 0$. Furthermore, $\theta(t)$ must be $0$ before reaching $X_f$, otherwise, it will grow exponentially fast as $r(t) \to 0$, which can be seen in equation (\ref{eq:r-deriv}).

@@

Now we will show that the fastest path between two points goes in a straight line. This will formalize the natural intuition that straight lines are the fastest path when there are no obstacles between the start and finish position.

\begin{theorem}
  Given a 2-dimensional particle $p$ whose initial position is $\bvec{X_1}$ and moves with bounded tangential acceleration, $a_t \le a_{t,max}$, and infinite centripetal acceleration, the fastest path $\hat{\gamma}(t)$ which $p$ can trace from $\bvec{X_1}$ to $\bvec{X_2}$ lies on the line segment from $\bvec{X_1}$ to $\bvec{X_2}$:
\end{theorem}
